---
output: github_document
---

```{r, include = FALSE}
library("dplyr")
library("readr")

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# LibreSense

Sensory analysis application.

This is the first attempt at building open-source software for sensory analysis.
In this first step, we are trying to build an application for Quantitative Descriptive Analysis (QDA).

We presented this project at LatinR 2018 in Buenos Aires, Argentina.

[http://47jaiio.sadio.org.ar/sites/default/files/LatinR_9.pdf](http://47jaiio.sadio.org.ar/sites/default/files/LatinR_9.pdf) [ES]

[https://www.youtube.com/watch?v=DqrVVRA7riA&feature=youtu.be](https://www.youtube.com/watch?v=DqrVVRA7riA&feature=youtu.be) [ES]

LibreSense contains two applications:

* The application for capturing panelist data -`run_panel`- the leader of the sensory panel can modify the name and number of the samples and also the descriptors.
* The results visualization application -`run_board`-.

In case you want to collaborate, check the issues.

## Installation

You can install the development version of `{libresense}` from [GitHub](https://github.com/) with:

``` r
if (!require("remotes")) {
  install.packages("remotes")
}
remotes::install_github(
  "anibalacatania/LibreSense", subdir = "libresense", dependencies = TRUE
)
```

## Example

### Setting the Attributes to Evaluate

To set which attributes -and their type- to evaluate by the panelists, this should be performed in a `csv` file, as exemplified in [atributos.csv](atributos.csv).

```{r, echo=FALSE}
read_csv("atributos.csv", col_types = cols()) %>% 
  as.data.frame() %>% 
  head()
```

This `csv` file should contain two columns:

* Nombre: The names of the attribute to be evaluated.
* Valores: The types of the attribute to be evaluated, must be one of `{"Numeric", "Text"}`

### Setting the Products to Evaluate

To set which products to evaluate by the panelists, this should be performed in a `csv` file, as exemplified in [productos.csv](productos.csv).

```{r, echo=FALSE}
read_csv("productos.csv", col_types = cols()) %>% 
  as.data.frame() %>% 
  head()
```

This `csv` file should contain at least one column, the name of the first column will be the name of the product to be evaluated, for example `"Copa"`. The values of this first column, will be the products to be evaluated. Additional columns will not be used by `{libresense}`, but might be useful for the panel leader.

### Running the App

#### Running the Panel

Let's assume we are using the two files presented above, to run the panel app, from an R console type:

```{r, eval=FALSE}
libresense::run_panel(
  products_file = "productos.csv",
  attributes_file = "atributos.csv",
  answers_dir = "Answers/"
)
```

#### Running the Board

Let's assume we are using the two files presented above, to run the board app, from an R console type:

```{r, eval=FALSE}
libresense::run_board(answers_dir = "Answers/")
```
